### 2017-04-27 Paper/paper.tex
It works! Still some minor horizontal phasing I belive... but overall yay! :)

\documentclass[12pt]{article}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{url}

\newcommand{\fft}{\text{FFT}}

\title{Audio Transport}
\date{\today}
\author{Trevor Henderson}

\begin{document}
\maketitle

\begin{abstract}
  yoyoyo.
\end{abstract}

\section{Introduction}

\section{Related Work}
Vocoders existed in the 70's.
vocoders take a monophonic source and spread it out over a range of pitches. requires harmonic similarity between the sources. Voices have lots of noise so it is easy - Daft punk.

realtime pitch shifting has become very popular in software and hardware. since the early 2000's.
Traditional pitch shifters were granular.
Updates to algorithms use phase vocoders.

Real time pitch shifting is a key element to autotune, which charachterized the music of the mid 2000's and is still popular today.

Harmonizers sound similar to vocoders. Take a pitch and pitch shift it into new chords.

In 2008 Melodyne pitch shifts individual elements without distoring the sound.
They must be processed beforehand and then can be manipulated in realtime

The 2012 product can manipulate polyphonic audio in realtime.
http://www.zynaptiq.com/pitchmap/

Most of this work has been proprietary and out of research.

So far as I can

processing but no synthesis.
\url{http://www.maths.lu.se/fileadmin/maths/personal_staff/Andreas_Jakobsson/ElvanderAKJ16_icassp_final.pdf}

http://remi.flamary.com/biblio/flamary2016optimal.pdf

\section{Optimal Transport}

Discussion of optimal transport.

Distance.

It also allows for you to interpolate between the two shapes along the trajectory.

In 1 dimension it is a solved problem.

For any dimensionality higher than 1, fast computation of the wasserstein2 distance is

Optimal transport is however only defined for, nonnegative masses.
Audio has phase.
Initial attempts to tackle the problem of phase involved formulating the transport problem differently.
But you can't hear the distance between two sinusoids that are played at the same pitch but at different phases
But you can hear the ``distance'' between two pitches.

\section{Phase Vocoders}
As noted by, 
audio quality is .

Vertical and horizontal incoherence.

\section{Algorithm Overview}

\subsection{STFT}
Weighted overlap add
root hann window
Let $M$ be the window size, $R$ be the hopsize.

We are guaranteed perfect reconstruction in the absense of spectral modifications if
\begin{align*}
  \sum_m w^2(n - mR) = 1,\;\forall n\in\mathbb{Z}
\end{align*}
I choose the to use the root Hann window.
\begin{align*}
  w(n) = 
  \cos(\pi n / M), 
  n\in
  \left[
  -\frac{M-1}{2}, 
  \frac{M-1}{2}
\right]
\end{align*}

\subsection{Spectrum Segmentation}
In order to fix the spec
Techniques mention peak finding.
However peak finding does not give a great sense of where to put boundaries. 
The reassigned frequency can be computed by:
\begin{align}
  \hat{\omega}(\omega) = 
  \omega + 
  \Im\left\{%
    \frac{X_\mathcal{D}(\omega)\cdot X^*(\omega)}{|X(\omega)|^2}
  \right\}
\end{align}
where $X_\mathcal{D}$ is the spectrum of the $x$ with the window
$w_\mathcal{D}(n) = \frac{d}{dt}w(n)$:
\begin{align}
  X_\mathcal{D}(\omega) = \fft\left(w_\mathcal{D}(n)x(n)\right)
\end{align}
d(root hann window)
\begin{align*}
   w_\mathcal{D}(n) = -\frac{\pi f_s}{M}\sin\left(\frac{\pi n}{M}\right)
\end{align*}

Nice plot.

\subsection{Phase Accumulation}
Horizontal coherence.
\begin{align*}
  \theta' = \theta + \frac{M}{f_s}\hat{\omega} 
\end{align*}

\subsection{Transport}

Combinations of frequencies:
rotating in the same direction
\begin{align}
  X_{01}(\omega) = 
  X_0(\omega)^{1-k}
  X_1(\omega)^k
\end{align}
This maintains that the energy of the signal is constant and that the phase rotates from one state to the other.
However, this rotation can happen in multiple directions:

\section{Implementation}
Real time.
\texttt{fftw}

\texttt{portaudio}
\texttt{portmidi}

video

github link

\section{Conclusion}

\section{References}

Fundemental theory

that phase vocoder one

1d transport paper?

STFT

weighted overlap add

fftw

portaudio

portmidi

cite justin solomon

\end{document}

### 2017-04-27 include/Vocoder/BidirectionalTransportVocoder.hpp

@@ -10,17 +10,23 @@
 class BidirectionalTransportVocoder : public PhaseVocoder {
   private:
     int sampleRate;
+    int windowSize;
+    int hopSize;
+
     double GROUP_DELAY_D_THRESHOLD = 100000000;
     double * amplitudes0;
     double * amplitudes1;
     double * phases0;
     double * phases1;
+    double * accumulatedPhases0;
+    double * accumulatedPhases1;
 
     SpectralMass * masses0;
     SpectralMass * masses1;
     SpectralMass * massesOut;
 
-    std::complex<double> * massesData;
+    double * massesAmpData;
+    double * massesPhaseData;
 
     std::size_t * assignmentIndices0;
     std::size_t * assignmentIndices1;
@@ -34,6 +40,8 @@ class BidirectionalTransportVocoder : public PhaseVocoder {
 
     std::size_t segementIntoMasses(
         std::complex<double> ** transforms,
+        double * amplitudes,
+        double * accumulatedPhases,
         SpectralMass * masses
         );
 
@@ -50,6 +58,12 @@ class BidirectionalTransportVocoder : public PhaseVocoder {
         std::complex<double> ** transforms,
         SpectralMass * masses,
         std::size_t numMasses);
+
+    void accumulatePhases(
+        const double * phases,
+        std::complex<double> ** transforms,
+        double * accumulatedPhases
+        );
 
   public:
     BidirectionalTransportVocoder(unsigned int hopSize, unsigned int windowSize, unsigned int sampleRate);

### 2017-04-27 include/Vocoder/SpectralMass.hpp

@@ -8,7 +8,8 @@ struct SpectralMass {
   std::size_t centerOfMass; // relative to start bin
   std::size_t leftLength;
   std::size_t rightLength;
-  std::complex<double> * centerOfMassPointer;
+  double * centerOfMassAmp;
+  double * centerOfMassPhase;
   double mass;
 };
 

### 2017-04-27 include/Vocoder/Transport.hpp

@@ -74,9 +74,9 @@ class Transport {
         double interpolationFactor);
 
     inline
-    static std::complex<double> rotate(
-        std::complex<double> value0,
-        std::complex<double> value1,
+    static double rotate(
+        double phase0,
+        double phase1,
         double interpolationFactor);
 };
 

### 2017-04-27 src/BidirectionalTransportVocoder.cpp

@@ -9,11 +9,12 @@
 #include "Vocoder/Transport.hpp"
 
 BidirectionalTransportVocoder::BidirectionalTransportVocoder(
-   unsigned int hopSize, 
+   unsigned int hopSize_, 
    unsigned int windowSize_,
    unsigned int sampleRate_
    ) : 
-  PhaseVocoder(hopSize, windowSize_, 1, sampleRate_),
+  PhaseVocoder(hopSize_, windowSize_, 1, sampleRate_),
+  hopSize(hopSize_),
   windowSize(windowSize_),
   sampleRate(sampleRate_)
 {
@@ -24,6 +25,10 @@ BidirectionalTransportVocoder::BidirectionalTransportVocoder(
 
   accumulatedPhases0 = new double[transformSize];
   accumulatedPhases1 = new double[transformSize];
+  for (std::size_t i = 0; i < transformSize; i++) {
+    accumulatedPhases0[i] = 0;
+    accumulatedPhases1[i] = 0;
+  }
 
   masses0 = new SpectralMass[transformSize];
   masses1 = new SpectralMass[transformSize];
@@ -33,7 +38,8 @@ BidirectionalTransportVocoder::BidirectionalTransportVocoder(
   // in the transform
   int maxTransportSize = 2*transformSize - 1;
 
-  massesData = new std::complex<double>[maxTransportSize];
+  massesAmpData = new double[maxTransportSize];
+  massesPhaseData = new double[maxTransportSize];
 
   assignmentIndices0 = new std::size_t[maxTransportSize];
   assignmentIndices1 = new std::size_t[maxTransportSize];
@@ -45,9 +51,13 @@ BidirectionalTransportVocoder::~BidirectionalTransportVocoder() {
   delete [] amplitudes1;
   delete [] phases0;
   delete [] phases1;
+  delete [] accumulatedPhases0;
+  delete [] accumulatedPhases1;
   delete [] masses0;
   delete [] masses1;
   delete [] massesOut;
+  delete [] massesAmpData;
+  delete [] massesPhaseData;
   delete [] assignmentIndices0;
   delete [] assignmentIndices1;
   delete [] assignmentMasses;
@@ -75,13 +85,27 @@ void BidirectionalTransportVocoder::processFrameTransform(
       );
 
   // Accumulate phases
+  accumulatePhases(
+      phases0,
+      transforms0,
+      accumulatedPhases0
+      );
+  accumulatePhases(
+      phases1,
[…]

### 2017-04-27 src/Transport.cpp

@@ -193,7 +193,8 @@ void Transport::transport<SpectralMass>(
     const std::size_t numMasses
     ) {
 
-  std::complex<double> * current = massesOut[0].centerOfMassPointer;
+  double * currentAmp = massesOut[0].centerOfMassAmp;
+  double * currentPhase = massesOut[0].centerOfMassPhase;
 
   for (std::size_t assignmentIndex = 0; assignmentIndex < numAssignments; assignmentIndex++) {
     std::size_t massIndex0 = assignmentIndices0[assignmentIndex];
@@ -219,55 +220,71 @@ void Transport::transport<SpectralMass>(
     massesOut[assignmentIndex].centerOfMass = centerOfMass;
     massesOut[assignmentIndex].leftLength = leftLength;
     massesOut[assignmentIndex].rightLength = rightLength;
-    massesOut[assignmentIndex].centerOfMassPointer = current + leftLength;
+    massesOut[assignmentIndex].centerOfMassAmp = currentAmp + leftLength;
+    massesOut[assignmentIndex].centerOfMassPhase = currentPhase + leftLength;
 
-    //double peakPhase = 
-      //mass0.centerOfMassPhase[0] * (1 - interpolationFactor) +
-      //mass1.centerOfMassPhase[0] * interpolationFactor;
+    // rotate from peak0 to peak1
+    double peakPhase = rotate(
+        mass0.centerOfMassPhase[0],
+        mass1.centerOfMassPhase[0],
+        interpolationFactor);
+
+    double newPeakPhase = 
+      mass0.centerOfMassPhase[0] * (1 - interpolationFactor) +
+      mass1.centerOfMassPhase[0] * interpolationFactor;
+
+    double phaseAdjustment = newPeakPhase - peakPhase;
 
     for (std::size_t i = 0; i < totalLength; i++) {
       long offset = i - leftLength;
 
-      //double amp0 = 
+      // Scale the amplitude
+      double amp0 = (mass0.centerOfMassAmp + offset)[0];
+      double amp1 = (mass1.centerOfMassAmp + offset)[0];
 
-      std::complex<double> data0 = (mass0.centerOfMassPointer + offset)[0];
-      std::complex<double> data1 = (mass1.centerOfMassPointer + offset)[0];
+      double ampOut = 
+        std::pow(scale0 * amp0, 1 - interpolationFactor) *
+        std::pow(scale1 * amp1, interpolationFactor);
+
+      // Rotate from phase0 to phase1
[…]

### 2017-04-27 test/BidirectionalTransportVocoderTest.cpp

@@ -11,7 +11,9 @@
 #include <Vocoder/BidirectionalTransportVocoder.hpp>
 
 int main() {
-  int hopSize = 4096;
+  //int hopSize = 4096;
+  int hopSize = 2048;
+  //int hopSize = 1024;
   int windowSize = 4096;
   int sampleRate = 44100;
   int numOutputChannels = 2;

### 2017-05-14 Paper/paper.tex
Added phase unwrapping, didn't really work. What if I just linearly interpolated the spectras once they've moved?? They couldn't interfere if they're centered to the same space

@@
 \section{Introduction}
+
+\section{Contributions}
+
+I developed an algorithm that
+
+techniques to
+reconstruct an audio signal
+
+I use a new technique to segment the spectrum into smaller pieces.
+Phase reconstruction
+
+I also implemented the algorithm in real-time to
 
@@
 \subsection{STFT}
-Weighted overlap add
-root hann window
-Let $M$ be the window size, $R$ be the hopsize.
 
-We are guaranteed perfect reconstruction in the absense of spectral modifications if
-\begin{align*}
-  \sum_m w^2(n - mR) = 1,\;\forall n\in\mathbb{Z}
-\end{align*}
-I choose the to use the root Hann window.
+Like almost all spectral algorithms, this one begins with a short-time Fourier transform (STFT) and ends with it's inverse (ISTFT).
+At the cost of reduced temporal resolution, the STFT gives us access to the spectral contents of the signal in time.
+
+The STFT fourier transform performs a discrete time Fourier transforms on windows of size $M$ of the original signal $x(n)$. 
+These windows are separated by a hop size of $R$. The $m$th frame of the STFT is
+
+\begin{align}
+  X_m(\omega) = 
+  \sum_{n = -\frac{M-1}{2}}^{\frac{M - 1}{2}} 
+  x_m(n)w(n)e^{-i\omega n}
+\end{align}
+
+$x_m(n) = x(n + mR)$ 
+is a buffer of size $M$ of the original signal
+shifted $R$ samples from the previous buffer.
+$w(n)$ is the synthesis window of size $M$.
+
+The ISTFT can then reconstruct the
+original signal by
+
+In the absence of spectral modifications, 
+we can achieve perfect reconstruction 
+of a signal given that the synthesis and analysis windows obey
+\begin{align}
+  \sum_m w(n - mR)f(n - mR) = 1,\; \forall n\in \mathbb{Z}
+\end{align}
+
+I chose to use the root Hann window, which gives perfect reconstruction when $R = \frac{M}{2k}$ for $k\in\mathbb{Z}$.
 \begin{align*}
@@
 \subsection{Spectrum Segmentation}
+
+In order to resolve the vertical incoherence --- the phasing issues within the frame --- my solution like others is to lock regions of the spectrum where phase relations are important.
+Rather than performing a transport over all bins, bins which are used to represent the same spectral information are lumped together.
+
+TODO
+Many applications use simple peak finding, which is good in most appliccations but fails in certain settings.
+Instead I determine where to split the window based on the \emph{reassigned frequency}.
+The reas
+
 In order to fix the spec
@@
 \section{Implementation}
+
+I implemented the above algorithm in real-time as an audio effect in C++.
+The program listens to two streams of audio
+which can be sent to it from a any diginal audio workstation, like Ableton Live.
+It then produces a new stream of audio that is interpolated between the two original streams according to a MIDI value.
+
 Real time.

### 2017-05-14 AudioUtilities/src/ISTFT.cpp

@@ -22,10 +22,13 @@ ISTFT::ISTFT(unsigned int hopSize_, unsigned int windowSize_) :
 
   // Calculate a window
   Window::rootHannWindow(synthesisWindow, windowSize);
-  
+
   // Clear the audio buffer
   for (int i = 0; i < windowSize; i++) {
     audioBuffer[i] = 0;
+
+    // TODO: delete
+    // synthesisWindow[i] = 1;
   }
 
   // Make a plan!

### 2017-05-14 AudioUtilities/src/Window.cpp

@@ -20,6 +20,9 @@ void Window::rootHannWindow(double * window, int windowSize) {
     // Center the window
     int windowIndex = i - halfWindow;
     window[i] = std::cos(M_PI * windowIndex/double(windowSize));
+
+    // TODO: delete
+    //window[i] *= window[i];
   }
 }
 
@@ -31,6 +34,9 @@ void Window::rootHannWindowD(double * window, int windowSize, int sampleRate) {
     // Center the window
     int windowIndex = i - halfWindow;
     window[i] = -(M_PI * sampleRate)/double(windowSize) * std::sin(M_PI * windowIndex/double(windowSize));
+
+    // TODO: delete
+    // window[i] = -(M_PI * sampleRate)/double(windowSize) * std::sin(2 * M_PI * windowIndex/double(windowSize));
   }
 
 }

### 2017-05-14 include/Vocoder/BidirectionalTransportVocoder.hpp

@@ -12,8 +12,8 @@ class BidirectionalTransportVocoder : public PhaseVocoder {
     int sampleRate;
     int windowSize;
     int hopSize;
+    double freqThreshold;
 
-    double GROUP_DELAY_D_THRESHOLD = 100000000;
     double * amplitudes0;
     double * amplitudes1;
     double * phases0;
@@ -38,7 +38,7 @@ class BidirectionalTransportVocoder : public PhaseVocoder {
         std::complex<double> * transform
         );
 
-    std::size_t segementIntoMasses(
+    std::size_t segmentIntoMasses(
         std::complex<double> ** transforms,
         double * amplitudes,
         double * accumulatedPhases,

### 2017-05-14 include/Vocoder/Transport.hpp

@@ -78,6 +78,11 @@ class Transport {
         double phase0,
         double phase1,
         double interpolationFactor);
+
+    inline
+    static double unwrap(
+        double phase,
+        double previousPhase);
 };
 
 #include "../../src/Transport.cpp"

### 2017-05-14 src/BidirectionalTransportVocoder.cpp

@@ -18,6 +18,7 @@ BidirectionalTransportVocoder::BidirectionalTransportVocoder(
   windowSize(windowSize_),
   sampleRate(sampleRate_)
 {
+  freqThreshold = sampleRate/double(windowSize);
   amplitudes0 = new double[transformSize];
   amplitudes1 = new double[transformSize];
   phases0 = new double[transformSize];
@@ -70,6 +71,8 @@ void BidirectionalTransportVocoder::processFrameTransform(
     std::complex<double> * transformOut
     ) {
 
+  // interpolationFactor = 0.5;
+
   // Convert to polar
   CoordinateTransforms::cartesianToPolar(
       transforms0[0],
@@ -97,12 +100,12 @@ void BidirectionalTransportVocoder::processFrameTransform(
       );
 
   // Segment masses
-  std::size_t numMasses0 = segementIntoMasses(
+  std::size_t numMasses0 = segmentIntoMasses(
       transforms0,
       amplitudes0,
       accumulatedPhases0,
       masses0);
-  std::size_t numMasses1 = segementIntoMasses(
+  std::size_t numMasses1 = segmentIntoMasses(
       transforms1,
       amplitudes1,
       accumulatedPhases1,
@@ -162,9 +165,41 @@ void BidirectionalTransportVocoder::processFrameTransform(
       transformOut
       );
 
+  //unsegment(
+      //masses0,
+      ////1,
+      //numMasses0,
+      //transformOut
+      //);
+
+  if (numMasses1 > 400 and numMasses1 > 400 and false) {
+    if (masses0[1].centerOfMass > 300 and masses1[1].centerOfMass > 300) {
+      SpectralMass mass = massesOut[0];
+
+      std::size_t totalLength = 1 + mass.leftLength + mass.rightLength;
+      std::cout << "Mass!: " << mass.mass << std::endl;
+      std::cout << "Total length!: " << totalLength << std::endl;
+      std::vector<double> amps(totalLength);
+      std::vector<double> phases(totalLength);
+
+      for (int i = 0 ; i < totalLength; i++) {
+        int offset = i - mass.leftLength;
+        amps[i] = (mass.centerOfMassAmp + offset)[0];
+        phases[i] = (mass.centerOfMassPhase + offset)[0] - 
+          (mass.centerOfMassPhase - mass.leftLength)[0];
+      }
+
+      std::vector<std::vector<double> > vectors = {
+        phases,
+        amps
+      };
+
+      Plotting::plotVectors(vectors);
+    }
+  }
[…]

### 2017-05-14 src/Transport.cpp

@@ -5,6 +5,8 @@
 #include <cmath>
 #include <complex>
 
+#include <iostream>
+
 #include "Vocoder/Transport.hpp"
 #include "Vocoder/SpectralMass.hpp"
 
@@ -223,18 +225,8 @@ void Transport::transport<SpectralMass>(
     massesOut[assignmentIndex].centerOfMassAmp = currentAmp + leftLength;
     massesOut[assignmentIndex].centerOfMassPhase = currentPhase + leftLength;
 
-    // rotate from peak0 to peak1
-    double peakPhase = rotate(
-        mass0.centerOfMassPhase[0],
-        mass1.centerOfMassPhase[0],
-        interpolationFactor);
-
-    double newPeakPhase = 
-      mass0.centerOfMassPhase[0] * (1 - interpolationFactor) +
-      mass1.centerOfMassPhase[0] * interpolationFactor;
-
-    double phaseAdjustment = newPeakPhase - peakPhase;
-
+    double prevPhase0 = (mass0.centerOfMassPhase - leftLength)[0];
+    double prevPhase1 = (mass1.centerOfMassPhase - leftLength)[0];
     for (std::size_t i = 0; i < totalLength; i++) {
       long offset = i - leftLength;
 
@@ -246,29 +238,50 @@ void Transport::transport<SpectralMass>(
         std::pow(scale0 * amp0, 1 - interpolationFactor) *
         std::pow(scale1 * amp1, interpolationFactor);
 
-      // Rotate from phase0 to phase1
+      // Unwrap phase
       double phase0 = (mass0.centerOfMassPhase + offset)[0];
       double phase1 = (mass1.centerOfMassPhase + offset)[0];
-      double phaseOut = rotate(
-          phase0,
-          phase1,
-          interpolationFactor);
+      double unwrappedPhase0 = unwrap(phase0, prevPhase0);
+      double unwrappedPhase1 = unwrap(phase1, prevPhase1);
+
+      // Average
+      double phaseOut = 
+        unwrappedPhase0 * (1 - interpolationFactor) + 
+        unwrappedPhase1 * interpolationFactor;
+      phaseOut = unwrappedPhase0;
       
-      // Rotate into peak frame
-      phaseOut += phaseAdjustment;
-
-      // The peak phase averages the accumulated phases
-
-      // Convolve with data that is
-      // scaled by assignment mass
       // and set to output
       (currentAmp + i)[0] = ampOut;
[…]

### 2017-05-14 test/BidirectionalTransportVocoderTest.cpp

@@ -12,12 +12,14 @@
 
 int main() {
   //int hopSize = 4096;
-  int hopSize = 2048;
-  //int hopSize = 1024;
+  int hopSize = 1024;
+  //int hopSize = 2048;
+  //int hopSize = 512;
+  //int windowSize = 2048;
   int windowSize = 4096;
   int sampleRate = 44100;
   int numOutputChannels = 2;
-  int numInputChannels = 4;
+  int numInputChannels = 6;
 
   int inputDeviceNum;
   int outputDeviceNum;
@@ -70,28 +72,30 @@ int main() {
 
     // Vocode
     leftVocoder.processFrame(
-        audio[0], 
         audio[2], 
+        audio[4], 
         interpolationFactor,
         output[0]
         );
     rightVocoder.processFrame(
-        audio[1], 
         audio[3], 
+        audio[5], 
         interpolationFactor,
         output[1]
         );
 
-    for (int channel = 0; channel < 2; channel ++) {
-      for (int i = 0; i < hopSize; i++) {
-        output[0][i] = 
-          output[0][i] * vocoderVolume + 
-          (audio[0][i] * (1 - interpolationFactor) + audio[2][i] *interpolationFactor) * linearVolume;
-        output[1][i] = 
-          output[1][i] * vocoderVolume + 
-          (audio[1][i] * (1 - interpolationFactor) + audio[3][i] *interpolationFactor) * linearVolume;
-      }
-    }
+    //for (int channel = 0; channel < 2; channel ++) {
+      //for (int i = 0; i < hopSize; i++) {
+        //output[0][i] = 
+          //output[0][i] * vocoderVolume + 
+          //(audio[0][i] * (1 - interpolationFactor) + audio[2][i] *interpolationFactor) * linearVolume +
+          //audio[4][i];
+        //output[1][i] = 
+          //output[1][i] * vocoderVolume + 
+          //(audio[1][i] * (1 - interpolationFactor) + audio[3][i] *interpolationFactor) * linearVolume +
+          //audio[5][i];
+      //}
+    //}
 
     outputStream.write(output);
   };

### 2017-05-14 test/BidirectionalVocoderTest.cpp

@@ -15,7 +15,7 @@ int main() {
   int windowSize = 4096;
   int sampleRate = 44100;
   int numOutputChannels = 2;
-  int numInputChannels = 4;
+  int numInputChannels = 6;
 
   int inputDeviceNum;
   int outputDeviceNum;
@@ -68,28 +68,28 @@ int main() {
 
     // Vocode
     leftVocoder.processFrame(
-        audio[0], 
         audio[2], 
+        audio[4], 
         interpolationFactor,
         output[0]
         );
     rightVocoder.processFrame(
-        audio[1], 
         audio[3], 
+        audio[5], 
         interpolationFactor,
         output[1]
         );
 
-    for (int channel = 0; channel < 2; channel ++) {
-      for (int i = 0; i < hopSize; i++) {
-        output[0][i] = 
-          output[0][i] * vocoderVolume + 
-          (audio[0][i] * (1 - interpolationFactor) + audio[2][i] *interpolationFactor) * linearVolume;
-        output[1][i] = 
-          output[1][i] * vocoderVolume + 
-          (audio[1][i] * (1 - interpolationFactor) + audio[3][i] *interpolationFactor) * linearVolume;
-      }
-    }
+    //for (int channel = 0; channel < 2; channel ++) {
+      //for (int i = 0; i < hopSize; i++) {
+        //output[0][i] = 
+          //output[0][i] * vocoderVolume + 
+          //(audio[0][i] * (1 - interpolationFactor) + audio[2][i] *interpolationFactor) * linearVolume;
+        //output[1][i] = 
+          //output[1][i] * vocoderVolume + 
+          //(audio[1][i] * (1 - interpolationFactor) + audio[3][i] *interpolationFactor) * linearVolume;
+      //}
+    //}
 
     outputStream.write(output);
   };

### 2017-05-14 src/BidirectionalTransportVocoder.cpp
IT WORRKKS WITH SAWSSS@@@

@@ -71,8 +71,6 @@ void BidirectionalTransportVocoder::processFrameTransform(
     std::complex<double> * transformOut
     ) {
 
-  // interpolationFactor = 0.5;
-
   // Convert to polar
   CoordinateTransforms::cartesianToPolar(
       transforms0[0],
@@ -129,6 +127,9 @@ void BidirectionalTransportVocoder::processFrameTransform(
       masses1
       );
 
+  //if (numMasses0 > 0) numMasses0 = 1;
+  //if (numMasses1 > 0) numMasses1 = 1;
+
   // Compute transformation matrix
   std::size_t numAssignments = 
     Transport::massAssignment<SpectralMass>(
@@ -166,9 +167,9 @@ void BidirectionalTransportVocoder::processFrameTransform(
       );
 
   //unsegment(
-      //masses0,
+      //masses1,
       ////1,
-      //numMasses0,
+      //numMasses1,
       //transformOut
       //);
 

### 2017-05-14 src/Transport.cpp

@@ -225,49 +225,39 @@ void Transport::transport<SpectralMass>(
     massesOut[assignmentIndex].centerOfMassAmp = currentAmp + leftLength;
     massesOut[assignmentIndex].centerOfMassPhase = currentPhase + leftLength;
 
-    double prevPhase0 = (mass0.centerOfMassPhase - leftLength)[0];
-    double prevPhase1 = (mass1.centerOfMassPhase - leftLength)[0];
-    for (std::size_t i = 0; i < totalLength; i++) {
-      long offset = i - leftLength;
-
-      // Scale the amplitude
-      double amp0 = (mass0.centerOfMassAmp + offset)[0];
-      double amp1 = (mass1.centerOfMassAmp + offset)[0];
-
-      double ampOut = 
-        std::pow(scale0 * amp0, 1 - interpolationFactor) *
-        std::pow(scale1 * amp1, interpolationFactor);
-
-      // Unwrap phase
-      double phase0 = (mass0.centerOfMassPhase + offset)[0];
-      double phase1 = (mass1.centerOfMassPhase + offset)[0];
-      double unwrappedPhase0 = unwrap(phase0, prevPhase0);
-      double unwrappedPhase1 = unwrap(phase1, prevPhase1);
-
-      // Average
-      double phaseOut = 
-        unwrappedPhase0 * (1 - interpolationFactor) + 
-        unwrappedPhase1 * interpolationFactor;
-      phaseOut = unwrappedPhase0;
-      
-      // and set to output
-      (currentAmp + i)[0] = ampOut;
-      (currentPhase + i)[0] = phaseOut;
-
-      prevPhase0 = unwrappedPhase0;
-      prevPhase1 = unwrappedPhase1;
-    }
-
     double desiredPeakPhase = 
       mass0.centerOfMassPhase[0] * (1 - interpolationFactor) +
       mass1.centerOfMassPhase[0] * interpolationFactor;
 
-    double peakPhase = (currentPhase + leftLength)[0];
-
-    double adjustment = desiredPeakPhase - peakPhase;
+    double phase0Adjustment = desiredPeakPhase - mass0.centerOfMassPhase[0];
+    double phase1Adjustment = desiredPeakPhase - mass1.centerOfMassPhase[0];
 
     for (std::size_t i = 0; i < totalLength; i++) {
-      (currentPhase + i)[0] += adjustment;
+      long offset = i - leftLength;
+
+      // Scale the amplitude
[…]

### 2017-05-14 test/BidirectionalTransportVocoderTest.cpp

@@ -11,8 +11,8 @@
 #include <Vocoder/BidirectionalTransportVocoder.hpp>
 
 int main() {
-  //int hopSize = 4096;
-  int hopSize = 1024;
+  int hopSize = 4096;
+  //int hopSize = 1024;
   //int hopSize = 2048;
   //int hopSize = 512;
   //int windowSize = 2048;

### 2017-05-14 include/Vocoder/Transport.hpp
Changed to slightly different structure

@@ -46,8 +46,8 @@ class Transport {
         const double * assignmentMasses,
         const std::size_t numAssignments,
         const double interpolationFactor,
-        MassContainer * massesOut,
-        const std::size_t numMasses
+        std::complex<double> * transformOut,
+        const std::size_t transformSize
         );
     
     template <class MassContainer>

### 2017-05-14 src/BidirectionalTransportVocoder.cpp

@@ -155,16 +155,16 @@ void BidirectionalTransportVocoder::processFrameTransform(
       assignmentMasses,
       numAssignments,
       interpolationFactor,
-      massesOut,
+      transformOut,
       transformSize
       );
 
   // Unsegment the masses
-  unsegment(
-      massesOut,
-      numAssignments,
-      transformOut
-      );
+  //unsegment(
+      //massesOut,
+      //numAssignments,
+      //transformOut
+      //);
 
   //unsegment(
       //masses1,

### 2017-05-14 src/Transport.cpp

@@ -144,39 +144,39 @@ std::size_t Transport::massAssignment(
   return numAssignments;
 }
 
-template <> inline
-void Transport::transport<double>(
-    const double * masses0,
-    const double * masses1,
-    const std::size_t * assignmentIndices0,
-    const std::size_t * assignmentIndices1,
-    const double * assignmentMasses,
-    const std::size_t numAssignments,
-    const double interpolationFactor,
-    double * massesOut,
-    const std::size_t numMasses
-    ) {
-  // Clear the current output
-  for (std::size_t i = 0; i < numMasses; i++) {
-    massesOut[i] = 0;
-  }
+//template <> inline
+//void Transport::transport<double>(
+    //const double * masses0,
+    //const double * masses1,
+    //const std::size_t * assignmentIndices0,
+    //const std::size_t * assignmentIndices1,
+    //const double * assignmentMasses,
+    //const std::size_t numAssignments,
+    //const double interpolationFactor,
+    //double * massesOut,
+    //const std::size_t numMasses
+    //) {
+  //// Clear the current output
+  //for (std::size_t i = 0; i < numMasses; i++) {
+    //massesOut[i] = 0;
+  //}
 
-  for (std::size_t assignmentIndex = 0; assignmentIndex < numAssignments; assignmentIndex++) {
-    std::size_t assignment0 = assignmentIndices0[assignmentIndex];
-    std::size_t assignment1 = assignmentIndices1[assignmentIndex];
+  //for (std::size_t assignmentIndex = 0; assignmentIndex < numAssignments; assignmentIndex++) {
+    //std::size_t assignment0 = assignmentIndices0[assignmentIndex];
+    //std::size_t assignment1 = assignmentIndices1[assignmentIndex];
 
-    double assignment = assignment0 * (1 - interpolationFactor) + assignment1 * interpolationFactor;
+    //double assignment = assignment0 * (1 - interpolationFactor) + assignment1 * interpolationFactor;
 
-    int leftAssignment = std::floor(assignment);
-    int rightAssignment = std::floor(assignment+1);
+    //int leftAssignment = std::floor(assignment);
+    //int rightAssignment = std::floor(assignment+1);
 
-    double leftContribution = rightAssignment - assignment;
-    double rightContribution = assignment - leftAssignment;
[…]

### 2017-05-14 test/BidirectionalTransportVocoderTest.cpp

@@ -11,8 +11,8 @@
 #include <Vocoder/BidirectionalTransportVocoder.hpp>
 
 int main() {
-  int hopSize = 4096;
-  //int hopSize = 1024;
+  //int hopSize = 4096;
+  int hopSize = 1024;
   //int hopSize = 2048;
   //int hopSize = 512;
   //int windowSize = 2048;

### 2017-05-15 src/Transport.cpp
Its done! Code is a mess but it works. Need to clean, write paper and make examples

@@ -214,8 +214,8 @@ void Transport::transport<SpectralMass>(
     double scale1 = assignmentMasses[assignmentIndex]/mass1.mass;
 
     // compute new data
-    std::size_t leftLength = std::min(mass0.leftLength, mass1.leftLength);
-    std::size_t rightLength = std::min(mass0.rightLength, mass1.rightLength);
+    std::size_t leftLength = std::max(mass0.leftLength, mass1.leftLength);
+    std::size_t rightLength = std::max(mass0.rightLength, mass1.rightLength);
 
     std::size_t totalLength = leftLength + rightLength + 1;
 
@@ -228,30 +228,25 @@ void Transport::transport<SpectralMass>(
     double phase0Adjustment = desiredPeakPhase - mass0.centerOfMassPhase[0];
     double phase1Adjustment = desiredPeakPhase - mass1.centerOfMassPhase[0];
 
-    for (std::size_t i = 0; i < totalLength; i++) {
-      long offset = i - leftLength;
-
-      // Scale the amplitude
+    for (int offset = -int(mass0.leftLength); offset <= int(mass0.rightLength); offset++) {
       double amp0 = scale0 * (mass0.centerOfMassAmp + offset)[0];
-      double amp1 = scale1 * (mass1.centerOfMassAmp + offset)[0];
-
       double phase0 = (mass0.centerOfMassPhase + offset)[0];
-      double phase1 = (mass1.centerOfMassPhase + offset)[0];
-
       phase0 += phase0Adjustment;
-      phase1 += phase1Adjustment;
 
-      std::complex<double> complex0 = std::polar(amp0, phase0);
-      std::complex<double> complex1 = std::polar(amp1, phase1);
-
-      std::complex<double> complexOut = 
-        complex0 * (1 - interpolationFactor) + 
-        complex1 * interpolationFactor;
-
-      // Amp out
       long outIndex = centerOfMass + offset;
       if (outIndex >=  0 and outIndex < transformSize) {
-        transformOut[outIndex] += complexOut;
+        transformOut[outIndex] += std::polar(amp0, phase0) * (1 - interpolationFactor);
+      }
+    }
+
+    for (int offset = -int(mass1.leftLength); offset <= int(mass1.rightLength); offset++) {
+      double amp1 = scale1 * (mass1.centerOfMassAmp + offset)[0];
+      double phase1 = (mass1.centerOfMassPhase + offset)[0];
[…]

### 2017-05-16 Paper/paper.tex
Updating notation in the paper. Becoming more complete

@@
 \usepackage{url}
+\usepackage{algpseudocode}
 
@@
 
+As we 
 In order to resolve the vertical incoherence --- the phasing issues within the frame --- my solution like others is to lock regions of the spectrum where phase relations are important.
@@
 
-\subsection{Phase Accumulation}
-Horizontal coherence.
-\begin{align*}
-  \theta' = \theta + \frac{M}{f_s}\hat{\omega} 
-\end{align*}
+$|S| \leq N$
+Segment into how at at most $N$ segments $m$ with arrays $X_m$ 
+with center of masses $c_m$ and masses $\rho_m$.
+\begin{align}
+  \sum_{m\in S} &\rho(m) = 1
+\end{align}
 
@@
 
-Combinations of frequencies:
-rotating in the same direction
+After two audio signals have been segmented into $S^0$ and $S^1$
+we then compute a transformation matrix
+$T:S^0\times S^1\to \mathbb{R}$
+that represents the transfer of mass from one segment to the other. 
+We want this matrix to minimize the distance the centers of each segment move.
+We do not want to transport a negative mass or more mass than is available:
 \begin{align}
-  X_{01}(\omega) = 
-  X_0(\omega)^{1-k}
-  X_1(\omega)^k
+  0 < T(m^0, m^1) \leq \max(\rho(m^0), \rho(m^1))
 \end{align}
-This maintains that the energy of the signal is constant and that the phase rotates from one state to the other.
-However, this rotation can happen in multiple directions:
+Additionally, we want to constrain that all of the mass receives an assignment:
+\begin{align}
+  \rho(m^0) = &\sum_{m^1\in S^1} T(m^0, m^1)
+  \;
+  \forall m^0\in S^0
+  \\
+  \rho(m^1) = &\sum_{m^0\in S^0} T(m^0, m^1)
+  \;
+  \forall m^1\in S^1
+\end{align}
+%We want to construct a set of assignments $T$, where each assignment $n\in T$ 
+%assigns a mass $\rho_n$ to be transported between segments $m_n^0$ and $m_n^1$.
+%\begin{align}
+  %0 < \rho_n \leq \max(\rho_{m_n^0}, \rho_{m_n^1})
+%\end{align}
+%Additionally, we want to constrain that all of the mass receives an assignment:
+%\begin{align}
+  %\rho_m = &\sum_{\substack{n\in T,\\m_n^0 = m}} \rho_n 
+  %\;
+  %\forall m\in S^0
+  %\\
+  %\rho_m = &\sum_{\substack{n\in T,\\m_n^1 = m}} \rho_n 
+  %\;
+  %\forall m\in S^1
+%\end{align}
+Given constraints 6, 7, 8 we want to find the assignment $T$ that minimizes the 2-wasserstein distance between the segments.
+\begin{align}
+  \min_T &\sum_{\substack{m^0\in S^0\\ m^1\in S^1}} T(m^0, m^1)|c(m^0) - c(m^1)|^2
+\end{align}
+We can satisfy the constrains and approximate the using the following algorithm:
+
+Plot of overlapping pieces
+
+Using this algorithm, we guarantee that the number of assignments is bounded by $|T| \leq 2N - 1$.
+
+\subsection{Phase Accumulation}
+
+When we move a segment of the spectrum to a new pitch, 
+the phases within that section rotate either slower or faster.
+This change does not make an audible difference within a window itself,
+but it can cause interference in the overlap between windows known as \emph{horizontal incoherence}.
+
+Include plot
+
+One solution to this in time-stretching is to estimate the phase of the next frame based on its new instantaneous frequency. 
+If a bin with phase $\theta_i^\tau$ is oscillating at exactly $\hat{\omega}_i$, the phase of that bin in the next window $\theta_i^{\tau+1}$ will be
+\begin{align*}
+  \theta(i, \tau+1) = 
+  \theta(i,\tau) + \frac{M}{f_s}\hat{\omega}
+\end{align*}
+This model is good for , but it breaks down at transients.
+Reinitialization steps.
+
+Fortunately however, we have at our disposal the phases of not one but two pitches.
+The center of  will oscillate at $(1 - k)\hat{\omega}(m_0) + \hat{\omega}^1_m$.
+Therefore we can interpolate.
+
+For every bin $i$ keep track of the accumulated phase $\theta_i^\tau$ which we want to have the following properties:
+
+\begin{align*}
+  \Theta_i^\tau &\approx
+  \sum_{t = 0}^\tau 
+  \frac{R}{f_s}\hat{\omega}_i^t\\
+  \Theta_i^\tau &= \theta_i^\tau\pmod{2\pi}
+\end{align*}
+
+Therefore we want the synthesis phase of $\theta_i$ at the center of mass of segment $m$ is
+\begin{align*}
+  \theta_{c_m} = (1-k)\Theta_{c_m^0} + k\Theta_{c_m^1}
+\end{align*}
+
+To preserve the vertical coherence between all bins within a singular
+we add $\theta_{c_m}' - \theta_{c_m}$ to all pitches
+\begin{align*}
+  \theta_i' = \theta_i + \theta_{c_m}' - \theta_{c_m}
+\end{align*}
+
+This way all the local relationships between phases are kept.
+
+\subsection{Resynthesis}
+
+Once the phases of each segment have been centered around the synthesis phase,
+we can synthesize a new spectrum:
+
+\begin{algorithmic}
+\\
+\For{$n\in\left[0,\frac{M-1}{2}\right]$}
+  \State{$Y(n) \leftarrow 0$}
+  \Comment{Clear the output}
+\EndFor
+\\
+\For{$m_0\in S_0$, $m_1\in S_1$}
+  \\
+  \State{$\hat{n} \leftarrow (1 - k)c(m_0) + kc(m_1)$}
+  \Comment{$\hat{n}$ is the new COM index}
+  \State{$\hat{\Theta} \leftarrow (1 - k)\Theta(c(m_0)) + k\Theta(c(m_1))$}
+  \Comment{$\theta'$ is the new COM phase}
+  \\
+  \For{$n\in U(m_0)$}
+    \\
+    \State{$n' \leftarrow n + \hat{n} - c(m_0)$}
+    \State{$\Theta' \leftarrow \Theta_0(n) + \hat{\Theta} - \Theta_0(c(m_0))$}
+    \\
+    \State{%
+      $Y(n') 
+      \leftarrow 
+      Y(n') +
+      (1 - k)
+      \frac{T(m_0, m_1)}{\rho(m_0)}
+      |X_0(n)|
+      e^{i\theta'}$
+    }
+    \\
+  \EndFor
+\EndFor
+\end{algorithmic}
+
+reentered by the phase accumulation, 
+we can simply linearly interpolate the audio and add it to the spectrum.
+Since the
 
@@
 \section{Conclusion}
+
+If one of the sources is time varying the source material quickly deteriorates.
+Great for constant sounds.
 

### 2017-05-18 Paper/paper.tex
More work on paper. Commiting for checkout

@@
 \usepackage{url}
+\usepackage{algorithm}
 \usepackage{algpseudocode}
@@
 \newcommand{\fft}{\text{FFT}}
+\newcommand{\ifft}{\text{IFFT}}
 
@@
 Like almost all spectral algorithms, this one begins with a short-time Fourier transform (STFT) and ends with it's inverse (ISTFT).
-At the cost of reduced temporal resolution, the STFT gives us access to the spectral contents of the signal in time.
+At the cost of reduced temporal resolution, the STFT gives us access to the spectral components of the signal in time.
 
-The STFT fourier transform performs a discrete time Fourier transforms on windows of size $M$ of the original signal $x(n)$. 
-These windows are separated by a hop size of $R$. The $m$th frame of the STFT is
+The STFT performs a discrete time Fourier transforms on windows of size $M$ of the original signal $x(n)$.
+These windows are separated by a hop size of $R$. The $\tau$th frame of the STFT is
+\begin{align}
+  X(\omega, \tau) = 
+  \text{DTFT}(\tilde{x}(n, \tau)w(n))
+\end{align}
+where
+\begin{align}
+  \tilde{x}(n,\tau) = x(n + \tau R)
+\end{align}
+is a buffer of the original signal, shifted $R$ samples from the previous buffer and $w(n)$ is the synthesis window.
+$w(n)$ is defined to be nonzero for $M$ samples, 
+$n\in\left[-\frac{M - 1}{2}, \frac{M - 1}{2}\right]$, so only the $M$ most recent samples of $x(n)$ need to be kept in memory.
 
+After some computation, 
+we produce frames $Y(\omega, \tau)$ 
+which we want to use to synthesize the signal $y(n)$.
+We perform this ISTFT using the weighted overlap add method.
+As the name suggests, this method works by
+overlapping adjacent frames of the IDTFT, 
+weighting them by an analysis window $f(n)$ 
+and then adding them together.
+The weighting helps to suppress audible discontinuities at frame boundaries
+which prevents artifacts in highly nonlinear filters such as the one this paper describes.
 \begin{align}
-  X_m(\omega) = 
-  \sum_{n = -\frac{M-1}{2}}^{\frac{M - 1}{2}} 
-  x_m(n)w(n)e^{-i\omega n}
+  \tilde{y}(n,\tau) &= \text{IDTFT}(Y(\omega, \tau))
+  \\
+  y(n) &= \sum_\tau f(n - \tau R)\tilde{y}(n - \tau R,\tau)
 \end{align}
 
-$x_m(n) = x(n + mR)$ 
-is a buffer of size $M$ of the original signal
-shifted $R$ samples from the previous buffer.
-$w(n)$ is the synthesis window of size $M$.
-
-The ISTFT can then reconstruct the
-original signal by
+Note that synthesizing a single sample of $y(n)$ requires overlapping
+$\frac{M}{R}$ frames which adds a latency of $M$ samples between $y$ and it's sources.
 
@@
 we can achieve perfect reconstruction 
-of a signal given that the synthesis and analysis windows obey
+of a signal transformed by the STFT and ISTFT given
+that the synthesis and analysis windows obey
 \begin{align}
-  \sum_m w(n - mR)f(n - mR) = 1,\; \forall n\in \mathbb{Z}
+  \sum_\tau w(n - \tau R)f(n - \tau R) = 1,\; \forall n\in \mathbb{Z}
+\end{align}
+I chose to use the square root of the Hann window as both the analysis and frequency window
+which gives perfect reconstruction when $R = \frac{M}{2k}$ for $k\in\mathbb{Z}$.
+\begin{align}
+  w(n) = 
+  f(n) =
+  \cos(\pi n / M)
+  %, 
+  %n\in
+  %\left[
+  %-\frac{M-1}{2}, 
+  %\frac{M-1}{2}
+%\right]
 \end{align}
 
-I chose to use the root Hann window, which gives perfect reconstruction when $R = \frac{M}{2k}$ for $k\in\mathbb{Z}$.
-\begin{align*}
-  w(n) = 
-  \cos(\pi n / M), 
-  n\in
-  \left[
-  -\frac{M-1}{2}, 
-  \frac{M-1}{2}
-\right]
-\end{align*}
+For actual computation we replace, the DTFT and IDTFT with the FFT and IFFT respectively. 
+The FFT has $M$ bins $n\in\left[-\frac{M - 1}{2}, \frac{M - 1}{2}\right]$.
+The frequency of the $n$th bin is given by
+\begin{align}
+  \omega = \frac{f_s n}{M}
+\end{align}
+where $f_s$ is sampling rate of the signal.
+Since the audio signal $x(n)$ is real, $X(\omega)$ is conjugate symmetric, so we only need to retain the positive bins $\left[0, \frac{M - 1}{2}\right]$ of the FFT.
 
@@
 \end{align}
-d(root hann window)
-\begin{align*}
+The derivative of our chosen window, the square root of the Hann window is:
+\begin{align}
    w_\mathcal{D}(n) = -\frac{\pi f_s}{M}\sin\left(\frac{\pi n}{M}\right)
-\end{align*}
+\end{align}
 
-Nice plot.
+TODO: Plot of segmentation
+TODO: Formalize segmentation algorithm
 
-$|S| \leq N$
-Segment into how at at most $N$ segments $m$ with arrays $X_m$ 
-with center of masses $c_m$ and masses $\rho_m$.
+This algorithm segments the spectrum into how at at most $N$ segments $s\in S$.
+Each is has a center $c(s)$ and a set of indices $U(s)$.
+The mass of each segment is given by 
 \begin{align}
-  \sum_{m\in S} &\rho(m) = 1
+  \rho(s) = \frac{\sum_{n\in U(s)}|X(n)|}{\sum_{n}|X(n)|}
+\end{align}
+Note that
+\begin{align}
+  \sum_{s\in S} \rho(s) = 1
 \end{align}
@@
 
-After two audio signals have been segmented into $S^0$ and $S^1$
+After two audio signals have been segmented into $S_0$ and $S_1$
 we then compute a transformation matrix
-$T:S^0\times S^1\to \mathbb{R}$
-that represents the transfer of mass from one segment to the other. 
-We want this matrix to minimize the distance the centers of each segment move.
+$T\in \mathbb{R}^{S_0\times S_1}$
+whose entries represents the transfer of mass from one segment to the other. 
+We want to construct this matrix so that the total movement of mass (or work) is minimized.
+%We want this matrix to minimize the distance the centers of each segment move.
 We do not want to transport a negative mass or more mass than is available:
 \begin{align}
-  0 < T(m^0, m^1) \leq \max(\rho(m^0), \rho(m^1))
+  0 < T(s_0, s_1) \leq \max(\rho(s_0), \rho(s_1))
 \end{align}
@@
 \begin{align}
-  \rho(m^0) = &\sum_{m^1\in S^1} T(m^0, m^1)
+  \rho(s_0) = &\sum_{s_1\in S_1} T(s_0, s_1)
   \;
-  \forall m^0\in S^0
+  \forall s_0\in S_0
   \\
-  \rho(m^1) = &\sum_{m^0\in S^0} T(m^0, m^1)
+  \rho(s_1) = &\sum_{s_0\in S_0} T(s_0, s_1)
   \;
-  \forall m^1\in S^1
+  \forall s_1\in S_1
 \end{align}
@@
 %\end{align}
-Given constraints 6, 7, 8 we want to find the assignment $T$ that minimizes the 2-wasserstein distance between the segments.
+Given these constraints we want to find the assignment $T$ that minimizes the 2-Wasserstein distance between the segments.
 \begin{align}
-  \min_T &\sum_{\substack{m^0\in S^0\\ m^1\in S^1}} T(m^0, m^1)|c(m^0) - c(m^1)|^2
+  \min_T &\sum_{\substack{s_0\in S^0\\ s_1\in S^1}} T(s_0, s_1)|c(s_0) - c(s_1)|^2
 \end{align}
-We can satisfy the constrains and approximate the using the following algorithm:
+We compute such a $T$ with algorithm~\ref{transport}.
+%We can satisfy the constrains and approximate the using the following algorithm:
 
-Plot of overlapping pieces
+\begin{algorithm}
+  \caption{Compute Transformation Matrix}\label{transport}
+\begin{algorithmic}
+  \State{$T \leftarrow 0$}
+  \State{$i,j \leftarrow 0$}
+  \State{$\rho_0 \leftarrow s_0^i$}
+  \State{$\rho_1 \leftarrow s_1^j$}
+  \While{$i < |S_0|$ and $j < |S_1|$}
+    \If{$\rho_0 < \rho_1$}
+      \State{$T(s_0^i, s_1^j) \leftarrow \rho_0$}
+      \State{$i \leftarrow i + 1$}
+      \State{$\rho_1 \leftarrow \rho_1 - \rho_0$}
+      \State{$\rho_0 \leftarrow \rho(s_0^i)$}
+    \Else
+      \State{$T(s_0^i, s_1^j) \leftarrow \rho_1$}
+      \State{$j \leftarrow j + 1$}
+      \State{$\rho_0 \leftarrow \rho_0 - \rho_1$}
+      \State{$\rho_1 \leftarrow \rho(s_1^i)$}
+    \EndIf
+  \EndWhile
+\end{algorithmic}
+\end{algorithm}
 
-Using this algorithm, we guarantee that the number of assignments is bounded by $|T| \leq 2N - 1$.
+TODO: Plot of overlapping pieces
 
+TODO: Proof - talk to Justin
+
+%Using this algorithm, the number of assignments is $2N - 1$.
+
+%From this point we could perform an interpolation.
+%With interpolation factor $k$ the new peak index would be
+%\begin{align}
+  %c' = (1 - k)c(s_0) + k(c(s_1))
+%\end{align}
+%However we must first deal with the phase
+\newpage
 \subsection{Phase Accumulation}
@@
 One solution to this in time-stretching is to estimate the phase of the next frame based on its new instantaneous frequency. 
-If a bin with phase $\theta_i^\tau$ is oscillating at exactly $\hat{\omega}_i$, the phase of that bin in the next window $\theta_i^{\tau+1}$ will be
-\begin{align*}
-  \theta(i, \tau+1) = 
-  \theta(i,\tau) + \frac{M}{f_s}\hat{\omega}
-\end{align*}
-This model is good for , but it breaks down at transients.
-Reinitialization steps.
+If a sinusoid with phase $\theta(\omega, \tau)$ 
+is oscillating with instantaneous frequency $\hat{\omega}(\omega)$ over frame $\tau$, we would expect the phase of the next frame to be
+\begin{align}
+  \theta(\omega, \tau+1) = 
+  \theta(\omega,\tau) + \frac{R}{f_s}\hat{\omega}(\omega)
+\end{align}
+This model is good for signals which do not undergo rapid changes in frequency. 
+However it does a poor job of dealing with transients in audio that can happen from sharp note attacks and percussive sounds. 
+To cope with this, people employ reinitialization steps in frames where transients are detected, however this method is.
 
@@
 
-\begin{align*}
+\begin{align}
   \Theta_i^\tau &\approx
@@
   \Theta_i^\tau &= \theta_i^\tau\pmod{2\pi}
-\end{align*}
+\end{align}
 
 Therefore we want the synthesis phase of $\theta_i$ at the center of mass of segment $m$ is
-\begin{align*}
+\begin{align}
   \theta_{c_m} = (1-k)\Theta_{c_m^0} + k\Theta_{c_m^1}
-\end{align*}
+\end{align}
 
@@
 we add $\theta_{c_m}' - \theta_{c_m}$ to all pitches
-\begin{align*}
+\begin{align}
   \theta_i' = \theta_i + \theta_{c_m}' - \theta_{c_m}
-\end{align*}
+\end{align}
 
@@
 \subsection{Resynthesis}
+
+Compute new center.
+Compute new phase for horizontal coherence
+Shift phases to preserve vertical coherence
+linearly interpolate between the resulting
 
@@
 \\
-\For{$n\in\left[0,\frac{M-1}{2}\right]$}
+\For{all $n$}
   \State{$Y(n) \leftarrow 0$}
@@
 which can be sent to it from a any diginal audio workstation, like Ableton Live.
-It then produces a new stream of audio that is interpolated between the two original streams according to a MIDI value.
+It then produces a new stream of audio 
+whose spectrum is transported between the two
+according to a MIDI value.
+The program depends on FFTW (Fastest Fourier Transform in the West) for Fourier transforms, and the open source libraries PortAudio and PortMidi for audio and MIDI handling respectively.
 
-Real time.
-\texttt{fftw}
+I work with a window size of 4096 samples and a sampling rate of 10 milliseconds. This is audible.
 
-\texttt{portaudio}
-\texttt{portmidi}
+My code is available at
+\url{https://github.com/sportdeath/Vocoder}
 
-video
-
-github link
+A video of my implimentation is available at.
 

### 2017-05-18 src/BidirectionalTransportVocoder.cpp

@@ -18,7 +18,7 @@ BidirectionalTransportVocoder::BidirectionalTransportVocoder(
   windowSize(windowSize_),
   sampleRate(sampleRate_)
 {
-  freqThreshold = sampleRate/double(windowSize);
+  freqThreshold = 1*sampleRate/double(windowSize);
   amplitudes0 = new double[transformSize];
   amplitudes1 = new double[transformSize];
   phases0 = new double[transformSize];

### 2017-05-18 test/BidirectionalTransportVocoderTest.cpp

@@ -11,15 +11,15 @@
 #include <Vocoder/BidirectionalTransportVocoder.hpp>
 
 int main() {
-  //int hopSize = 4096;
+  //int hopSize = 1024;
+  //int hopSize = 1024;
   int hopSize = 1024;
-  //int hopSize = 2048;
   //int hopSize = 512;
   //int windowSize = 2048;
   int windowSize = 4096;
   int sampleRate = 44100;
   int numOutputChannels = 2;
-  int numInputChannels = 6;
+  int numInputChannels = 8;
 
   int inputDeviceNum;
   int outputDeviceNum;
@@ -84,18 +84,11 @@ int main() {
         output[1]
         );
 
-    //for (int channel = 0; channel < 2; channel ++) {
-      //for (int i = 0; i < hopSize; i++) {
-        //output[0][i] = 
-          //output[0][i] * vocoderVolume + 
-          //(audio[0][i] * (1 - interpolationFactor) + audio[2][i] *interpolationFactor) * linearVolume +
-          //audio[4][i];
-        //output[1][i] = 
-          //output[1][i] * vocoderVolume + 
-          //(audio[1][i] * (1 - interpolationFactor) + audio[3][i] *interpolationFactor) * linearVolume +
-          //audio[5][i];
-      //}
-    //}
+    for (int channel = 0; channel < 2; channel ++) {
+      for (int i = 0; i < hopSize; i++) {
+        output[channel][i] += audio[channel + 6][i];
+      }
+    }
 
     outputStream.write(output);
   };

### 2017-05-18 paper/paper2.tex
Paper almost done

--- 2017-05-18 Paper/paper.tex
@@
 
-A video of my implimentation is available at.
+A video of my implimentation in action is available at.
 
@@
 
-If one of the sources is time varying the source material quickly deteriorates.
-Great for constant sounds.
+As seen in the implimentation video, the sound quality is good for many types of signals.
+
+One of the most noticeable artifacts that still exists is the interference that results from a single pitch being mapped to many places.
+As the note splits and moves to new locations it begins to interfere with itself.
+This can cause a drop in volume at the very start or end of a transport.
+This could be fixed by requiring that the mapping be fixed.
+
+There are however sounds that the effect does not work for.
+If we try to transport to audio which has rapid changes in pitch, the source material deteriorates quickly because the transport map is not consistent.
+In certain settings however this can be used musically.
+If for example the we are transporting between two sounds with relatively constant pitch.
+
+But overall this the sound quality is good for many sounds.
+It has the potential to be used by DJs looking for interesting ways to mix between songs, and musicians looking to create new sounds.
 

### 2017-05-18 src/BidirectionalTransportVocoder.cpp

@@ -18,7 +18,7 @@ BidirectionalTransportVocoder::BidirectionalTransportVocoder(
   windowSize(windowSize_),
   sampleRate(sampleRate_)
 {
-  freqThreshold = 1*sampleRate/double(windowSize);
+  freqThreshold = M_PI*sampleRate/double(windowSize);
   amplitudes0 = new double[transformSize];
   amplitudes1 = new double[transformSize];
   phases0 = new double[transformSize];

### 2017-05-18 Paper/paper.tex
Paper almost done

@@
 \begin{abstract}
-  yoyoyo.
+  This paper describes a new technique to interpolate between audio signals.
+  This technique uses optimal transport to move the pitches in one signal to the pitches in another.
+  Audibly this sounds like a generalized portamento, or glide between sounds.
+  The main contributions of this paper are methods to remove the different types of phasing distortion that can occur in a naive transport.
+  A real time implementation of the algorithm produces musical results.
 \end{abstract}
@@
 
+A portamento is musical term used to describe one pitch sliding into another.
+The portamento can be found in almost all genres of music, 
+whether it be a slur in a vocal passage or in the twang of a slide guitar.
+It provides smooth way to transition between notes and even chords.
+
+However many instruments 
+Most acoustic keyboard instruments like the piano or organ example can't glide at all. 
+This is fixed with a pitch bend or for monophic instruments with a glide knob.
+These systems are difficult to move between chords of different shapes.
+Moreover all of these methods require that the glide happen between sounds of the same timbre.
+
+This paper describes a method for protamento not only between sounds with different timbres but between sounds from different instruments entirely.
+It does this using optimal transport, a problem mostly found in computer graphics and machine learning applications. 
+It also relies on machinery which is common to phase vocoders.
+
+This paper aims to fix that problem
+
+It functions by
+
+And even instruments that can glide are often limited in the chord shapes they can glide between.
+
+In music the portamento is limi
+Few instruments for example can glide smoothly between chords of different shapes, 
+For example gliding between  chords of different shapes becomes challenging and gliding between chords with different
+
+This paper aims to develop a generalized port
+
+Portamentos can also be polyphonic. For example the slide guitar technique, a technique commonly found in blues and country music, is in essense a polyphonic portamento between chords on the guitar.
+This paper aims to develop 
+sounds that map between the pitches of audio file.
+
+for example uses this sound 
+The goal of this paper is to
+
 \section{Contributions}
 
-I developed an algorithm that
+I developed a new technique to interpolate between two audio signals using optimal transport.
+The optimal transport problem, discussed in detail later, is defined only for positive real inputs. 
+The spectra of an audio signal however is complex.
+It is natural that the transport be done on the amplitudes of these complex numbers, but this begs the question of what is to be done with the phase.
 
-techniques to
-reconstruct an audio signal
+This paper describes several types of distortion that can occur because of phasing issues and how to resolve them. 
+These problems are similar to those experienced in phase vocoders --- the technology that allows an audio's speed to be changed without changing the pitch.
+However the solutions posed in this paper improve upon the techniques used in phase vocoders and take advantage of the specifics of this problem.
 
-I use a new technique to segment the spectrum into smaller pieces.
-Phase reconstruction
-
-I also implemented the algorithm in real-time to
+The new techniques introduced in this paper are 
+a method to segment the spectrum into groups whose local phase relations have audible significance.
+A method to update the phases of a piece of the spectrum that has been moved to a new pitch.
+A method to synthesize the audio
 
 \section{Related Work}
-Vocoders existed in the 70's.
-vocoders take a monophonic source and spread it out over a range of pitches. requires harmonic similarity between the sources. Voices have lots of noise so it is easy - Daft punk.
 
-realtime pitch shifting has become very popular in software and hardware. since the early 2000's.
-Traditional pitch shifters were granular.
-Updates to algorithms use phase vocoders.
+Many audio tools exist that modulate the pitch of an audio signal.
+The phase vocoder, which became practical in the early 2000's is now pervasive in all types of music production. 
+The phase vocoder is also the technology that enables auto-tune --- an effect which swept through the music industry in the mid-2000's and is still popular today.
 
-Real time pitch shifting is a key element to autotune, which charachterized the music of the mid 2000's and is still popular today.
+Several audio tools work to manipulate the pitches within an audio signal seperately.
+Melodyne is a popular tool in the music industry for correcting the pitches of notes.
+In 2008, Melodyne released a feature to adjust the pitches of individual notes in polyphonic music, allowing a user to rewrite the harmony of a song.
+This technology works on a static audio file, but in 2012 a product named QQQQ
+allowed for individual pitch manipulation to happen in real time.
 
-Harmonizers sound similar to vocoders. Take a pitch and pitch shift it into new chords.
-
-In 2008 Melodyne pitch shifts individual elements without distoring the sound.
-They must be processed beforehand and then can be manipulated in realtime
-
-The 2012 product can manipulate polyphonic audio in realtime.
+However neither of these products pitch shift on with granularity that this method does.
+In addition they rely on the user to select notes and decide where they want to go.
+While this is an extremely useful feature in the studio, it is hard to control in the live setting.
+This method allows the user to simply specify where they would like the pitches to go with another audio file and they can go there.
 http://www.zynaptiq.com/pitchmap/
 
-Most of this work has been proprietary and out of research.
-
-So far as I can
-
-processing but no synthesis.
-\url{http://www.maths.lu.se/fileadmin/maths/personal_staff/Andreas_Jakobsson/ElvanderAKJ16_icassp_final.pdf}
-
-http://remi.flamary.com/biblio/flamary2016optimal.pdf
+Optimal transport has been used for audio processing before.
+However so far as I can tell, no paper has synthesized sound with optimal transport.
+Most of the existing papers in this cross section use optimal transport to help transcribe music.
+By using a distance metric based on the harmonic series seems to be useful for determining where a note exists.
 
@@
 \end{align}
-
 Note that synthesizing a single sample of $y(n)$ requires overlapping
@@
 
+As seen in the implimentation video, the sound quality is good for many types of signals.
+
+One of the most noticeable artifacts that still exists is the interference that results from a single pitch being mapped to many places.
+As the note splits and moves to new locations it begins to interfere with itself.
+This can cause a drop in volume at the very start or end of a transport.
+This could be fixed by requiring that the mapping be fixed.
+
+There are however sounds that the effect does not work for.
+If we try to transport to audio which has rapid changes in pitch, the source material deteriorates quickly because the transport map is not consistent.
+In certain settings however this can be used musically.
+If for example the we are transporting between two sounds with relatively constant pitch.
+
+But overall this the sound quality is good for many sounds.
+It has the potential to be used by DJs looking for interesting ways to mix between songs, and musicians looking to create new sounds.
+
 If one of the sources is time varying the source material quickly deteriorates.
@@
 \end{document}
+
+PLOT:
+optimal transport
+plot reassigned frequency
+plot segmentation
+plot buffered segmentation
+plot misaligned audio
+Phasing
+
+Why did I do this frequency reassignment method?
+
+It can pick peaks in the presence of noise, but no threshold so even small peaks can be detected.
+
+Doesn't require setting a peak threshold
+
+Then Write algorithm for frequency
+
+Clean up phase accumulation math
+
+What are the symbols in resynthesize for center and phase
+
+Then clean up some text
+
+Then do references
+
+Then do slides
+
+Then do Demo
+
+Get rid of optimal transport and phase vocoder sections
+Put description of optimal transport in the introduction.
+Intersperse Phase vocoder stuff

### 2017-05-22 paper/paper.tex
Clean up

\documentclass[12pt]{article}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{subcaption}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{cite}

\newcommand{\fft}{\text{FFT}}
\newcommand{\ifft}{\text{IFFT}}

\title{Audio Transport}
\date{\today}
\author{Trevor Henderson}

\begin{document}
\maketitle

\begin{abstract}
  This paper describes a new technique to interpolate between audio signals.
  This technique uses optimal transport to move the pitches in one signal to the pitches in another.
  Audibly this sounds like a generalized portamento, or glide between sounds.
  The main contributions of this paper are methods to remove the different types of phasing distortion that can occur in a naive transport.
  A real time implementation of the algorithm produces musical results.
  \\
  \\
  Code: \href{https://github.com/sportdeath/Vocoder}{https://github.com/sportdeath/Vocoder}.
  \\
  Video: \href{https://youtu.be/PQGV0fk3Gww}{https://youtu.be/PQGV0fk3Gww}
\end{abstract}

\section{Introduction}

A portamento is a musical term used to describe one pitch sliding into another.
The portamento can be found in almost all genres of music, 
whether it is a slur in a vocal passage or in the twang of a slide guitar.
It provides smooth way to transition between notes and even chords.

This effect, though widely used, is limited in its use by the physical design of the instrument. 
Most acoustic keyboard instruments for example, like the piano or organ, can't glide at all because notes are formed from a discrete set of strings or pipes. 
Digital keyboards can produce a glide with  a pitch wheel.
The pitch wheel can change the playback rate of a sample as it plays or use spectral time stretching to change the pitch of a prerecorded audio sample.
However, this system does not make it easy to slide between chords of different shapes.
And moreover, it is impossible to slide between sounds with different timbres.

This paper describes a method to produce a portamento not only between sounds with different timbres but between sounds from different instruments entirely.
It does this using optimal transport, a geometric technique typically found in applications to computer graphics and machine learning.

The optimal transport problem seeks a map between two probability masses in $\mathbb{R}^n$ that minimizes the work done on the mass.
Mass in the transport must physically move through space and all mass must be preserved.
\begin{figure}
  %{r}{0.65\textwidth}
  \begin{center}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim={0 27.5cm 0 0},clip,width=\textwidth]{figures/euclidean_gray}
    \caption{Linear Interpolation}
    \label{fig:linear}
  \end{subfigure}
  \hspace{1cm}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim={0 27.5cm 0 0},clip,width=\textwidth]{figures/wasserstein}
    \caption{Optimal Transport Interpolation}
    \label{fig:wasserstein}
  \end{subfigure}
  \end{center}
  \caption{From \emph{Convolutional Wasserstien Distances: Efficient Optimal Transportation on Geometric Domains} by Justin Solomon et al, 2015\cite{justin}}
  \label{fig:justin}
\end{figure}
I find that optimal transport is best explained visually as in figure~\ref{fig:justin} from Justin Solomon et al \cite{justin}.
Efficient solutions to compute an optimal transport mapping is an open problem 
for $n > 1$.
However, optimal transport on the real line has a closed form solution.

We will use optimal transport on the real line to find a mapping between the spectra of two audio sources that minimizes the distances that the pitches move. 
From this assumption we can guarantee that no pitches will cross each other.

\section{Contributions}

I developed a new technique to interpolate between two audio signals using optimal transport.
The optimal transport problem is usually only defined only for positive real inputs. 
The spectra of an audio signal however is complex.
It is natural that the transport be done on the absolute values of these complex numbers, but this begs the question of what is to be done with the phase.

This paper describes several types of distortion that can occur because of phasing issues and how to resolve them. 
These problems are similar to those experienced in phase vocoders --- the technology that allows an audio signal's speed to be changed without changing the pitch.
However the solutions posed in this paper improve upon the techniques used in phase vocoders and take advantage of the specifics of this problem set up.
\cite{vocoder}

These are the main techniques used in the algorithm which are new to this paper:
\begin{enumerate}
  \item
    A method to segment and audio spectrum into groups whose local phase relations have audible significance.
  \item
    A method to update the phase of a pitch shifted signal given a source and destination.
\end{enumerate}

Additionally we also provide a real-time implementation of the algorithm for audible demonstration.

\section{Related Work}

Many audio tools exist that modulate the pitch of an audio signal.
The phase vocoder, also given other names in application like `time-warping,' became practical in the early 2000's, is now pervasive in all types of music production\cite{vocoder}\cite{ableton}. It is useful to piece together samples played in different keys, or to fine tune a recording.
The phase vocoder is also the technology that enables auto-tune --- an effect which swept through the music industry in the mid-2000's and is still popular today.

Several audio tools work to manipulate the pitches within an audio signal separately.
Melodyne is a popular tool in the music industry for correcting the pitches of notes.
In 2008, Melodyne released a feature to adjust the pitches of individual notes in polyphonic music, allowing a user to rewrite the harmony of a song
\cite{melodyne}.
This technology works on a static audio file, but in 2012 a product named Zynaptiq allows for different bands of an audio stream to be pitch shifted
\cite{zynaptiq}.

However neither of these products pitch shift on with granularity that is needed to slide between notes of different timbres.
In addition, they rely on the user to select which notes they would like to tune.
While this is an extremely useful feature in the studio, it is hard to control in the live setting.

Optimal transport has been used for audio processing before.
However so far as I can tell, no paper has synthesized sound with optimal transport.
Most of the existing papers in this cross section use optimal transport to help transcribe music.
Using a distance metric based on the harmonic series seems to be useful for determining fundamental of a note\cite{musictransport1}\cite{musictransport2}.

\section{Algorithm Overview}

\subsection{STFT}

Like almost all spectral algorithms, this one begins with a short-time Fourier transform (STFT) and ends with it's inverse (ISTFT).
At the cost of reduced temporal resolution, the STFT gives us access to the spectral components of the signal in time.
The following summarizes the relavant properties of the STFT for this problem as described by Julius O. Smith III in \emph{Spectral Audio Signal Processing} \cite{dsprelated}.

The STFT performs a discrete time Fourier transforms on windows of size $M$ of the original signal $x(n)$.
These windows are separated by a hop size of $R$. The $\tau$th frame of the STFT is
\begin{align}
  X(\omega, \tau) = 
  \text{DTFT}(\tilde{x}(n, \tau)w(n))
\end{align}
where
\begin{align}
  \tilde{x}(n,\tau) = x(n + \tau R)
\end{align}
is a buffer of the original signal, shifted $R$ samples from the previous buffer and $w(n)$ is the synthesis window.
$w(n)$ is defined to be nonzero for $M$ samples, 
$n\in\left[-\frac{M - 1}{2}, \frac{M - 1}{2}\right]$, so only the $M$ most recent samples of $x(n)$ need to be kept in memory.

After some computation, 
we produce frames $Y(\omega, \tau)$ 
which we want to use to synthesize the signal $y(n)$.
We perform this ISTFT using the weighted overlap add method.
As the name suggests, this method works by
overlapping adjacent frames of the IDTFT, 
weighting them by an analysis window $f(n)$ 
and then adding them together.
The weighting helps to suppress audible discontinuities at frame boundaries
which prevents artifacts in highly nonlinear filters such as the one this paper describes.
\begin{align}
  \tilde{y}(n,\tau) &= \text{IDTFT}(Y(\omega, \tau))
  \\
  y(n) &= \sum_\tau f(n - \tau R)\tilde{y}(n - \tau R,\tau)
\end{align}
Note that synthesizing a single sample of $y(n)$ requires overlapping
$\frac{M}{R}$ frames which adds a latency of $M$ samples between $y$ and it's sources.

In the absence of spectral modifications, 
we can achieve perfect reconstruction 
of a signal transformed by the STFT and ISTFT given
that the synthesis and analysis windows obey
\begin{align}
  \sum_\tau w(n - \tau R)f(n - \tau R) = 1,\; \forall n\in \mathbb{Z}
\end{align}
I chose to use the square root of the Hann window as both the analysis and frequency window
which gives perfect reconstruction when $R = \frac{M}{2k}$ for $k\in\mathbb{Z}$.
\begin{align}
  w(n) = 
  f(n) =
  \cos(\pi n / M)
  %, 
  %n\in
  %\left[
  %-\frac{M-1}{2}, 
  %\frac{M-1}{2}
%\right]
\end{align}

For actual computation we replace, the DTFT and IDTFT with the FFT and IFFT respectively. 
The FFT has $M$ bins $n\in\left[-\frac{M - 1}{2}, \frac{M - 1}{2}\right]$.
The frequency of the $n$th bin is given by
\begin{align}
  \omega = 2\pi\frac{f_s n}{M}
\end{align}
where $f_s$ is sampling rate of the signal.
Since the audio signal $x(n)$ is real, $X(\omega)$ is conjugate symmetric, so we only need to retain the positive bins $\left[0, \frac{M - 1}{2}\right]$ of the FFT.

\subsection{Spectrum Segmentation}

\begin{figure}[!h]
  %{r}{0.65\textwidth}
  \begin{center}
  \begin{subfigure}[b]{0.65\textwidth}
    \includegraphics[width=\textwidth]{figures/VerticalIncoherence0}
    \caption{Interpolated Spectra}
    \label{fig:verticalincoherence0}
  \end{subfigure}
  \hspace{1cm}
  \begin{subfigure}[b]{0.65\textwidth}
    \includegraphics[width=\textwidth]{figures/VerticalIncoherence1}
    \caption{Interpolated Waveform}
    \label{fig:verticalincoherence1}
  \end{subfigure}
  \end{center}
  \caption{%
    If we perform a transport across all bins,
    sinusoids smeared across multiple bins can be split in two (\ref{fig:verticalincoherence0}).
    The resulting waveform 
    (\ref{fig:verticalincoherence1}) should the sum of two sinusoids, but it
    experiences distortion
    known as \emph{vertical incoherence}.
    %If we do optimal transport on the amplitudes, we end up splitting the .
    %The lower plot is the interpolation of sinusoids using this method.
    %It is obviously distorted with what is known as \emph{vertical incoherence} --- phasing distortion within a window of the STFT.
  }
  \label{fig:verticalincoherence}
\end{figure}

Once we have the spectrum $X_0$ from one window of the STFT, we then want to transport it to another spectrum $X_1$ from a separate STFT running in parallel.
However performing a transport across all bins results in distortion as seen in figure~\ref{fig:verticalincoherence}.
This distortion is known as vertical incoherence in phase vocoder algorithms. It occurs when local phase relations pertaining to a particular frequency component are lost
\cite{vocoder}.

A technique that is used to fix this problem in phase vocoders is to lock the phases of particular regions of the spectrum
\cite{vocoder}.
For this problem, this means we must do the transport on chunks of the spectrum rather than the individual bins.
All that is needed now is an algorithm to segment the spectrum into pieces whose phase relations we want to maintain.

Phase vocoder algorithms often use peak finding to determine these regions
\cite{vocoder}.
However, peak finding is not very robust to noise. 
Any small oscillation in the transform 
can cause an unnecessary split and produce vertical incoherence.
This does not seem to effect the performance of the phase vocoder because all of these regions are moving to the same pitch.
But the effect becomes much more noticeable when suddenly all regions are moving to entirely different pitches.
Adding a noise threshold not only makes the segmentation less elegant, it also begins to ignore high frequency components, which typically have a much smaller amplitude.

\begin{figure}[!h]
  %{r}{0.65\textwidth}
  \begin{center}
  \includegraphics[width=0.7\textwidth]{figures/ReassignedFreqCMaj}
  \caption{The reassigned frequencies in a C Major chord played on the piano}
  \label{fig:reassignedfrequency}
\end{center}
\end{figure}

Instead, I determine where to split the window based on the \emph{reassigned frequency}, $\hat{\omega}(\omega)$.
The reassigned frequency as shown in figure~\ref{fig:reassignedfrequency} 
is a more accurate notion of the frequency a bin is actually oscillating at.
A sinusoid smeared over multiple bins for example, should have a constant reassigned frequency, because the smeared amplitudes all support the center frequency.
Likewise, many of the peaks in the spectrum of the piano chord in figure~\ref{fig:reassignedfrequency} align with relatively flat sections of the reassigned frequency
\cite{reassignment}.

\begin{algorithm}
  \caption{Segment Spectrum}\label{segment}
\begin{algorithmic}
  \State{$S\leftarrow\{\}$}
  \State{$s\leftarrow\emptyset$}
  \For{each bin $n>0$}
    \State{$U(s) \leftarrow U(s)\cup n$}

    \\
    \If{%
      $\hat{\omega}(n - 1) - \omega(n - 1) > 0$ and
      $\hat{\omega}(n) - \omega(n) < 0$}
      \State{// \emph{$\hat{\omega}$ crosses down through $\omega$}}
      \State{// \emph{Make a center of mass}}
      \State{$c(s) \leftarrow n - 1$}
      \\
    \ElsIf{%
      $\hat{\omega}(n - 1) - \omega(n - 1) < 0$ and
      $\hat{\omega}(n) - \omega(n) > 0$}
      \State{// \emph{$\hat{\omega}$ crosses up through $\omega$}}
      \State{// \emph{Split the mass}}
      \State{$S\leftarrow S\cup\{s\}$}
      \State{$s\leftarrow\emptyset$}
    \EndIf
  \EndFor
\end{algorithmic}
\end{algorithm}
Once the reassigned frequency has been computed, we can use algorithm~\ref{segment} to segment the spectrum into a set $S$ of spectral pieces $s(i)$.
Each piece $s$ is defined by its center of mass $c(s)$, its set of bins $U(s)$, and a mass $\rho(s)$.
The center of the mass is picked so that it is the bin whose frequency is closest to the reassigned frequency.
The mass of each segment is a normalized sum of its components
\begin{align}
  \rho(s) = \frac{\sum_{n\in U(s)}|X(n)|}{\sum_{n}|X(n)|}
\end{align}
This normalization occurs so that the sum of masses equals 1 for optimal transport:
\begin{align}
  \sum_{s\in S} \rho(s) = 1
\end{align}

\begin{figure}[!h]
  %{r}{0.65\textwidth}
  \begin{center}
  \includegraphics[width=0.7\textwidth]{figures/SegmentedReassignedFreqCMaj}
  \caption{The segmentation produced by algorithm~\ref{segment}. Centers of mass are marked in green and the boundaries between masses are marked in black.}
  \label{fig:segmentation}
\end{center}
\end{figure}

Figure~\ref{fig:segmentation} shows a segmentation performed by the algorithm.
The figure shows that many of the center of mass marked in green align exactly with the peak. 
However the small peak near $\omega = 8000 \;\text{radians}\cdot \text{s}^{-1}$ is ignored because it is dominated by the larger peak next to it which still exerts influence.

\begin{figure}[!h]
  %{r}{0.65\textwidth}
  \begin{center}
  \includegraphics[width=0.7\textwidth]{figures/ReassignedFreqChri1}
  \caption{The reassigned frequency of a linear chirp --- a sinusoid with linearly increasing amplitude. The reassigned frequency is approximately linear over the chirps range. The dotted black lines show a region where small fluctuations do not produce new segmentations.}
    \label{fig:chirp}
  \end{center}
\end{figure}

A slight modification must be made to this algorithm to account for audio where the relation between frequency and reassigned frequency is relatively linear as shown in figure~\ref{fig:chirp}. In these cases, slight perturbations in the reassigned frequency with cause it to fluctuate over and under the bin frequency producing unnecessary segmentations.
We resolve this by adding a boundary region whose width is equal to the width of an FFT bin, $2\pi \frac{f_s}{M}$.
The algorithm is essentially the same only we form masses when we cross up through the entire boundary region
and define our center of masses when we cross down through the entire boundary region.

The reassigned frequency can be computed as follows \cite{reassignment}:
\begin{align}
  \hat{\omega}(\omega) = 
  \omega + 
  \Im\left\{%
    \frac{X_\mathcal{D}(\omega)\cdot X^*(\omega)}{|X(\omega)|^2}
  \right\}
\end{align}
where $X_\mathcal{D}$ is the spectrum of the input signal $x(n)$ windowed with
$w_\mathcal{D}(n) = \frac{d}{dt}w(n)$:
\begin{align}
  X_\mathcal{D}(\omega) = \fft\left(w_\mathcal{D}(n)x(n)\right)
\end{align}
The derivative of our chosen window, the square root of the Hann window is:
\begin{align}
   w_\mathcal{D}(n) = -\frac{\pi f_s}{M}\sin\left(\frac{\pi n}{M}\right)
\end{align}

\subsection{Transport}

After two audio signals have been segmented into $S_0$ and $S_1$
we then compute a transformation matrix
$T\in \mathbb{R}^{|S_0|\times |S_1|}$
whose entries represents the transfer of mass from one segment to the other. 
We want to construct this matrix so that the total movement of mass (or work) is minimized.
%We want this matrix to minimize the distance the centers of each segment move.
We do not want to transport a negative mass or more mass than is available:
\begin{align}
  0 < T(i, j) \leq \max(\rho(s_0(i)), \rho(s_1(j)))
\end{align}
Additionally, we want to constrain that all of the mass receives an assignment:
\begin{align}
  \rho(s_0(i)) = &\sum_{j} T(i, j)
  \;
  \forall i
  \\
  \rho(s_1(j)) = &\sum_{i} T(i, j)
  \;
  \forall j
\end{align}
%We want to construct a set of assignments $T$, where each assignment $n\in T$ 
%assigns a mass $\rho_n$ to be transported between segments $m_n^0$ and $m_n^1$.
%\begin{align}
  %0 < \rho_n \leq \max(\rho_{m_n^0}, \rho_{m_n^1})
%\end{align}
%Additionally, we want to constrain that all of the mass receives an assignment:
%\begin{align}
  %\rho_m = &\sum_{\substack{n\in T,\\m_n^0 = m}} \rho_n 
  %\;
  %\forall m\in S^0
  %\\
  %\rho_m = &\sum_{\substack{n\in T,\\m_n^1 = m}} \rho_n 
  %\;
  %\forall m\in S^1
%\end{align}
Given these constraints we want to find the assignment $T$ that minimizes the 2-Wasserstein distance between the segments
\cite{justin}.
\begin{align}
  \min_T &\sum_{i,j} T(i, j)|c(s_0(i)) - c(s_1(j))|^2
\end{align}

\begin{algorithm}
  \caption{Compute Transformation Matrix}\label{transport}
\begin{algorithmic}
  \State{$T \leftarrow 0$}
  \State{$i,j \leftarrow 0$}
  \State{$\rho_0 \leftarrow s_0(i)$}
  \State{$\rho_1 \leftarrow s_1(j)$}
  \While{$i < |S_0|$ and $j < |S_1|$}
    \If{$\rho_0 < \rho_1$}
      \State{$T(i, j) \leftarrow \rho_0$}
      \State{$i \leftarrow i + 1$}
      \State{$\rho_1 \leftarrow \rho_1 - \rho_0$}
      \State{$\rho_0 \leftarrow \rho(s_0(i))$}
    \Else
      \State{$T(i, j) \leftarrow \rho_1$}
      \State{$j \leftarrow j + 1$}
      \State{$\rho_0 \leftarrow \rho_0 - \rho_1$}
      \State{$\rho_1 \leftarrow \rho(s_1(j))$}
    \EndIf
  \EndWhile
\end{algorithmic}
\end{algorithm}

We compute such a $T$ with algorithm~\ref{transport}.
%We can satisfy the constrains and approximate the using the following algorithm:
This algorithm takes advantage of the fact that in a minimized mapping, no mass passes over any other mass.
The leftmost unmapped mass in one segmentation must always come from the leftmost unmapped mass in the other.
So we can greedily assign masses from left to right.

\begin{figure}[!h]
  %{r}{0.65\textwidth}
  \begin{center}
  \includegraphics[width=0.7\textwidth]{figures/TransportDiagram}
  \caption{A visualization of the 1D optimal transport algorithm.
  The width of the bins represents the amount of mass in that bin.}
  \label{fig:optimaltransport}
\end{center}
\end{figure}

This algorithm runs linearly and guarantees that $T$ will be sparse with no more than $2N - 1$ entries.
This algorithm is visualized in figure~\ref{fig:optimaltransport}

%Using this algorithm, the number of assignments is $2N - 1$.

%From this point we could perform an interpolation.
%With interpolation factor $k$ the new peak index would be
%\begin{align}
  %c' = (1 - k)c(s_0) + k(c(s_1))
%\end{align}
%However we must first deal with the phase
\subsection{Phase Accumulation}

\begin{figure}[!h]
  %{r}{0.65\textwidth}
  \begin{center}
  \includegraphics[width=0.7\textwidth]{figures/HorizontalIncoherence0}
  \caption{%
    The pitch of the original signal in blue is increased to be that of the orange signal.
  The window boundary at $t = 0$ experiences horizontal incoherence.}
  \label{fig:horizontalincoherence}
\end{center}
\end{figure}

After we have a transport map, we need to actually repitch the spectral segments.
When we move a segment of the spectrum to a new pitch, 
the phases within that section rotate either slower or faster.
This change does not make an audible difference within a window itself,
but it can cause interference in the overlap between windows known as \emph{horizontal incoherence} seen in figure~\ref{fig:horizontalincoherence}.

One solution to this in the realm of phase vocoders is to estimate the phase of the next frame based on its instantaneous frequency
\cite{vocoder}.
If a sinusoid with phase $\angle X(\omega, \tau)$ 
is oscillating with instantaneous frequency $\hat{\omega}(\omega)$ over frame $\tau$, we would expect the phase of the next frame to be
\begin{align}
  \angle X(\omega, \tau+1) = 
  \angle X(\omega,\tau) + \frac{R}{f_s}\hat{\omega}(\omega)
\end{align}
This model is good for signals which do not undergo rapid changes in frequency. 
However it does a poor job of dealing with transients in audio that can happen from sharp note attacks and percussive sounds. 
To cope with this, some phase vocoders employ phase reinitialization steps in frames where transients are detected.
Even with this step additional phasing can still occur
\cite{vocoder}.

Fortunately however, we have at our disposal the phases of not one but two pitches.
If we are interpolating along the transport map with interpolation factor $k$,
then for a particular assignment between $s_0$ and $s_1$
we know the center frequency of the mass will oscillate at frequency:
\begin{align}
  \hat{\omega}' = (1 - k)\hat{\omega}(c(s_0)) + \hat{\omega}(c(s_1))
\end{align}
The accumulated phases 
$\Theta_0(\omega, \tau)$
and
$\Theta_1(\omega, \tau)$
each change proportional to $\omega_0$
and $\omega_1$.
So to get a new phase $\Theta'$ whose rate of change is proportional to $\hat{\omega}'$, we can linearly interpolate between $\Theta_0$ and $\Theta_1$.
\begin{align}
  \Theta' = (1 - k)\Theta_0(\omega, \tau) + k\Theta_1(\omega, \tau)
\end{align}
We calculate the accumulated phase so that it increases with respect to $\hat{\omega}$, but round it to maintain the exact phase. This rounding is essentially a reinitialization every step, which we get away with because of the two sources.
\begin{align}
  \Theta(\omega, \tau) &\approx
  \sum_{t = 0}^\tau 
  \frac{R}{f_s}\hat{\omega}(\omega, \tau)\\
  \Theta(\omega, \tau) &= \Theta(\omega,\tau)\pmod{2\pi}
\end{align}

\subsection{Resynthesis}

\begin{algorithm}
  \caption{Resynthesize}\label{resynthesize}
\begin{algorithmic}
\\
\For{all $n$}
  \State{$Y(n) \leftarrow 0$}
  \Comment{Clear the output}
\EndFor
\\
\For{$i$, $j$}
  \\
  \State{$c' \leftarrow (1 - k)c(s_0(i)) + kc(s_1(j))$}
  \Comment{$c'$ is the new center index}
  \State{$\Theta' \leftarrow (1 - k)\Theta(c(s_0(i))) + k\Theta(c(s_1(j)))$}
  \Comment{$\Theta'$ is the new center phase}
  \\
  \For{$n\in U(s_0(i))$}
    \\
    \State{$n' \leftarrow n + \hat{n} - c(s_0(i))$}
    \Comment{$n'$ is the new bin index}
    \State{$\theta' \leftarrow \Theta_0(n) + \Theta' - \Theta_0(c(s_0(i)))$}
    \Comment{$\Theta'$ is the new bin phase}
    \\
    \State{%
      $Y(n') 
      \leftarrow 
      Y(n') +
      (1 - k)
      \frac{T(i, j)}{\rho(s_0(i))}
      |X_0(n)|
      e^{i\theta'}$
    }
    \Comment{Linearly interpolate}
    \\
  \EndFor

  \For{$n\in U(s_1(j))$}
    \\
    \State{$n' \leftarrow n + \hat{n} - c(s_1(j))$}
    \Comment{$n'$ is the new bin index}
    \State{$\theta' \leftarrow \Theta_1(n) + \Theta' - \Theta_1(c(s_1(j)))$}
    \Comment{$\Theta'$ is the new bin phase}
    \\
    \State{%
      $Y(n') 
      \leftarrow 
      Y(n') +
      j
      \frac{T(i, j)}{\rho(s_1(j))}
      |X_1(n)|
      e^{i\theta'}$
    }
    \Comment{Linearly interpolate}
    \\
  \EndFor
\EndFor
\end{algorithmic}
\end{algorithm}

After we have segmented, transported, and accumulated the phases, we can finally synthesize an output fourier transform $Y(n)$ as in algorithm~\ref{resynthesize}.
For every assignment, we find the new center frequency, and new center phase.
We then adjust all the phases within the segment to maintain local phase relations.
Finally we linearly interpolate between all of the segments according to the interpolation factor.

\section{Implementation}

I implemented the above algorithm in real-time as an audio effect in C++.
The program listens to two streams of audio
which can be sent to it from a any digital audio workstation.
It then produces a new stream of audio 
whose spectrum is transported between the two
according to a MIDI value.
The program depends on the library FFTW (Fastest Fourier Transform in the West) \cite{fftw} for Fourier transforms, and the open source libraries PortAudio\cite{portaudio} and PortMidi\cite{portmidi} for audio and MIDI handling respectively.

I work with a window size of 4096 samples and a sampling rate of 44,100 hertz.
This produces a delay of 10 milliseconds which is audible.
However the delay only exists on from the perspective of the source audio.
There is no perceivable delay from the perspective of the midi controller 
because the MIDI value is read right before a frame is synthesized.

Code: \href{https://github.com/sportdeath/Vocoder}{https://github.com/sportdeath/Vocoder}.

Video:
\href{https://youtu.be/PQGV0fk3Gww}{https://youtu.be/PQGV0fk3Gww}

\section{Conclusion}

As seen in the implementation video, the algorithm produces undistorted signals for a variety of inputs: monophonic - polyphonic, acoustic  - synthesized, bright - dark.

One of the most noticeable artifacts that still exists is the interference that results from a single pitch being mapped to many places.
As the note splits and moves to new locations it interferes with itself for the brief period of time where the splits overlap.
This artifact causes a drop in volume at the very start or end of a transport.
This type of interference has plagued music far before computers.
If adjacent semi-tones on a piano are played together you can hear beating.
As of now, I have no idea how to fix that other than performing a transport that
only maps a frequency to a single other frequency.
But that takes some of the fun away.
And in the end the distortion doesn't sound harsh and digital and in many cases the volume dip is barely noticeable.

Overall the effect produced a variety of interesting sounds.
This effect has potential both in the studio and live setting. It can be used by DJs to transition between songs or could be used by guitarists to swap between dualing solos.
It functions as a new way to filter between sounds of different timbres.
Moreover some of the sounds it produces sound simply unexplainable and I'm excited to explore their possibilities.

\nocite{*}
\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}

Fundemental theory

that phase vocoder one

1d transport paper?

STFT

weighted overlap add

fftw

portaudio

portmidi

cite justin solomon

PLOT:
optimal transport
plot reassigned frequency
plot segmentation
plot buffered segmentation
plot misaligned audio
Phasing

Why did I do this frequency reassignment method?

It can pick peaks in the presence of noise, but no threshold so even small peaks can be detected.

Doesn't require setting a peak threshold

Then Write algorithm for frequency

Clean up phase accumulation math

What are the symbols in resynthesize for center and phase

Then clean up some text

Then do references

Then do slides

Then do Demo

Get rid of optimal transport and phase vocoder sections
Put description of optimal transport in the introduction.
Intersperse Phase vocoder stuff
